%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -*- Mode: Latex -*- %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% uhtest-appendix.tex -- 
%% Author          : Robert Brewer
%% Created On      : Fri Oct  2 16:31:12 1998
%% Last Modified By: Robert Brewer
%% Last Modified On: Mon Oct  5 14:41:05 1998
%% RCS: $Id: uhtest-appendix.tex,v 1.1 1998/10/06 02:07:03 rbrewer Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Copyright (C) 1998 Robert Brewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 

\chapter{Table of Notations}
\noindent\begin{tabularx}{\linewidth}{@{}>{\hsize=.4\hsize}X>{\hsize=1.5\hsize}X@{}}
	
	$A$ & Bathymetry Grid - A 2D Grid of  non-negative real numbers that describe the depth of a section of the sea floor.  $A_{i,j}$ refers to the value in the cell at row i, column j.  $A_{x}$ refers to the value in cell $x$.\\

	$B$ & Behavior Grid - A 2D Grid of  non-negative real numbers that indicate the percentage of all transmissions that are expected to be released from that cell over the entire experiment.  $B_{i,j}$ refers to the value in the cell at row i, column j.  $B_{x}$ refers to the value in cell $x$.\\
	
	$G$ & Goodness Grid - A 2D Grid of real numbers with the same dimensions as the Bathymetry Grid.  Contains real numbers that indicate how good a particular cell will serve as a receiver placement.  $G_{i,j}$ refers to the value in the cell at row i, column j.  $G_{x}$ refers to the value in cell $x$.\\

	$n$ & The square root of the total number of cells in the Bathymetry Grid (A).  Approximates the edge dimension of a square Bathymetry Grid.\\

	$k$ & The fixed elevation of receivers off of the sea floor in meters.  Assumed to be constant for all receiver emplacements in the simulation.\\
	
	$ERT_{x,T,i,j}$ & Estimated Receivable Transmissions - The number of transmissions that are expected to be received from a cell at row $i$, column $j$ by a receiver at cell $T$, according to Evaluation Algorithm $x$.\\

	$P_{Attenuation}(x)$ & Probability of Detection due to attenuation - Describes the probability of detecting an acoustic transmission originating at a range of $x$ meters away from a receiver.\\

	$P_{observation}(T,a,b)$ & Percentage of transmission in cell ($a,b$) which are observable (due to bathymetric obstruction) from cell $T$ .  Given a receiver in cell $T$, and some distribution of animals in the water column of cell ($a,b$), this value represents the percentage of observable animals in ($a,b$) from A.  Animal distribution is handled by the Vertical Movement Model (Section~\ref{verticalMovement}).  Section~\ref{LoSAlgorithm} gives the algorithm for determining bathymetric obstruction and Line of Sight.\\
	
	$Range(T,i,j)$ &  Distance from a particular receiver location $T$ to the cell at row i, column j.  Computed as the Euclidean distance between the center of a receiver-containing cell $T$ and the center of the cell at row i, column j on a 2D Grid.\\
	
	$R_{opt}$ & The number of optimal receivers to be placed by the system (excluding projected receivers). \\
	
	$R_{proj}$ & The number of projected receivers to be placed by the system.\\
	
	$D_max(p, q)$& The deepest observable depth within a target cell $q$, relative to the receiver containing cell $p$. See Section~\ref{LoSAlgorithm}\\
	



\end{tabularx}

\noindent\begin{tabularx}{\linewidth}{@{}>{\hsize=.3\hsize}X>{\hsize=1.5\hsize}X@{}}
	$Edist_{p,v}$ & The Euclidean distance between cells $p$ and $v$.\\

	$m_{p,q}$&The slope between cells $p$ and $q$, expressed as the change in bathymetric elevation (of the sea floor) between cells $p$ and $q$, divided by the Euclidean Distance between cells $p$ and $q$. See Section~\ref{LoSAlgorithm}\\
	
	$D_{range}$ & The detection range of the receiver.  Assumed to be the same for all receivers in the network.  Given as the distance between a receiver and tag at which there is a 5\% chance of detecting an acoustic transmission.\\
	
	$d_{detection}$& The edge size of the Detection Area, equal to $2r+1$ cells.\\
	$f_{supp}$ & The Suppression Range Factor.  A positive real number.\\
	$r_{supp}$ & The Suppression Range. Equal to $f_{supp}*D_{range}$\\
	$d_{supp}$ & The square root of the number of cells in the Suppression Area.  The edge dimension of the Suppression Area.  Equal to $2*r_{supp} + 1$ cells.\\
\end{tabularx}